### 2017-06-15 paper/paper.tex
Updated paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  My documentation report
%  Objetive: Explain what I did and how, so someone can continue with the investigation
%
% Important note:
% Chapter heading images should have a 2:1 width:height ratio,
% e.g. 920px width and 460px height.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Filmchat}

%----------------------------------------------------------------------------------------
% PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt,fleqn]{book} % Default font size and left-justified equations

\usepackage[top=3cm,bottom=3cm,left=3.2cm,right=3.2cm,headsep=10pt,letterpaper]{geometry} % Page margins

\usepackage{xcolor} % Required for specifying colors by name
\definecolor{ocre}{RGB}{52,177,201} % Define the orange color used for highlighting throughout the book

% Font Settings
\usepackage{avant} % Use the Avantgarde font for headings
%\usepackage{times} % Use the Times font for headings
\usepackage{mathptmx} % Use the Adobe Times Roman as the default text font together with math symbols from the Sym­bol, Chancery and Com­puter Modern fonts

\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[utf8]{inputenc} % Required for including letters with accents
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs

% Bibliography
%\usepackage[style=alphabetic,sorting=nyt,sortcites=true,autopunct=true,babel=hyphen,hyperref=true,abbreviate=false,backref=true,backend=biber]{biblatex}

%\defbibheading{bibempty}{}

\usepackage[style=ieee,sorting=nyt,sortcites=true,autopunct=true,babel=hyphen,hyperref=true,abbreviate=false,backref=true,backend=biber]{biblatex}
\addbibresource{Mendeley_MA-230.bib} % BibTeX bibliography file
\defbibheading{bibempty}{}

\input{structure} % Insert the commands.tex file which contains the majority of the structure behind the template

\begin{document}

%----------------------------------------------------------------------------------------
% TITLE PAGE
%----------------------------------------------------------------------------------------

\begingroup
\thispagestyle{empty}
\AddToShipoutPicture*{\put(0,0){\includegraphics[scale=1.25]{frontpage3}}} % Image background
\centering
\vspace*{5cm}
\par\normalfont\fontsize{32}{32}\sffamily\selectfont
\textbf{Filmchat}\\
{\LARGE Movie recommendation as NUI}\par % Book title
\vspace*{1cm}
{\Huge Tarjei Utnes}\par % Author name
\endgroup

%----------------------------------------------------------------------------------------
% TABLE OF CONTENTS
%----------------------------------------------------------------------------------------

\chapterimage{head3.png} % Table of contents heading image

\pagestyle{empty} % No headers

\tableofcontents % Print the table of contents itself

%\cleardoublepage % Forces the first chapter to start on an odd page so it's on the right

\pagestyle{fancy} % Print headers again

%----------------------------------------------------------------------------------------
% CHAPTER 1 Project Description
%----------------------------------------------------------------------------------------

\chapterimage{head3.png} % Chapter heading image

\chapter{Project Description}
\section{Vision}
Imagine the following scenario.You, and one or more friends, or maybe a romantic partner, are sitting in the sofa in front of the TV. You are having a movie night, but you are not quite sure what you should see. While you sit and discuss your likes, and dislikes, there is a microphone hidden in the corner. The listens to what you are saying, and it is sending it to an Interactive System. It understands, who is speaking, who is liking a movie, and who is not liking a movie. It also has, from earlier experience, knowledge of what you and maybe the ones you are with have seen, and liked. It also knows what you did not like.

After a few minutes discussing with the people it starts rolling some suggestions over the television.

By using your voice, you can tell the Interactive system to play a trailer, if it is showing you something you think might be interesting.

When you have decided it knows on what streaming, or maybe even traditional linear TV the movie plays on and selects that option for you.

The name of the system, is Filmchat.

The goal of the system; To help you decide on the best movie for you, and your friends to watch right now.

This project is aimed for use at home, or possibly on-line in chat rooms or forums. The goal of the project is to interpret and come up with suggestions for which movies, or TV series its users should watch. I have limited the scope in this project to implement parts of the project, because of the available time given.
There are competing or projects existing that are close to doing what this project is aiming to do. Below is a non-exhaustive list of such projects, and solutions.
\section{Competing and nearby Interactive Systems}\index{competing systems}
While the list is by no means an exhaustive one, these are some of the known interactive systems doing things near to what I want my system to do. None of them has the complete solution.
\subsection{“And Chill” Netflix recommendation bot}\index{AndChill}
The “And Chill” chatbot:
\begin{quote}
Well, that’s when the Facebook chatbot And Chill comes in handy. The bot asks you to suggest a film you like with a few reasons why, or to simply explain what you’re looking for in your next movie.  
Within a few moments, the bot should issue you with a short selection of trailers. It sometimes takes a while to compute, but it’s still a pretty efficient way of narrowing down Netflix’s massive library.
That said, if it’s running at peak capacity, you’ll have to wait your turn before it takes your requests. A representative for the app told a Facebook commenter last month that they were working on scaling capacity.\cite{Williams2016TheUK}
\end{quote}
While the And Chill bot, has a nice, based on my experience, recommendation system it is still the task of having to actually sit and chat to a bot. This, to me feels unnecessarily tedious, and somewhat unnatural. However the recommendation engine that sits behind the bot I think is very good. 
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{and_chill_bot.png}
  \caption{And Chill recommending a movie}
  \label{fig:and_chill}
\end{figure}

\subsection{Filmgrail}\index{Filmgrail}
Filmgrail is an app to check where you are able to watch the movie or television series you have selected. What makes Filmgrail unique in a Norwegian context, is that it covers many of the Norwegian streaming services. \cite{AmundsenAppenAftenposten}.
At the moment, however, it does not have a "recommendation" engine. It has some "Popular searches" and other similar things, but nothing that can recommend based on what you have watched, or what you like. This could change in the future.
\subsection{Google Assistant}\index{googleassistant}
Google Assistant can also recommend movies. See Fig \ref{fig:google_assistant}. It uses my geographical location, or context to recommend a movie that I am able to watch on the cinema nearby. In the example shown in Fig \ref{fig:google_assistant} I was using speech recognition to have it recommend the movie.

The recommendations I have gotten from the Google Assistan has not been very exciting, and to me atleast, not very relevant. The speech recognition however, is very good. Better than the one I was able to implement using Googles Speech API.

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{google_assistant.png}
  \caption{Google Assitant recommending a movie based on geographic context}
  \label{fig:google_assistant}
\end{figure}

%----------------------------------------------------------------------------------------
% CHAPTER 2 Interactive System Description
%----------------------------------------------------------------------------------------
\chapterimage{head3.png}

\chapter{Interactive System: Description}
There are three parts implemented in this system
The first part listens to a given channel on Slack, using the outgoing web hook\cite{SlackOutgoingSlack}.This is the TextBlob Botlisener.Then there is the component that listens to a post message and tries to find a movie from that message. This is the Omdbsearcher.
The third part is the Speech Regognizer which is heavily based on the Csharp example from Googles Speech API \cite{GoogleCloudPlatformSpeechPlatform} and listens to the microphone. It then tries to search for the terms in the recognized speech and if it finds a movie within it it posts it to the same Slack channel.

\section{TextBlob Botlistener}
A Python web service that uses Textblob\cite{TextBlobTutorial:Documentation} as a way of getting noun phrase, and sentiment from the text it is provided. It gets input from the designated slack channel, using the outgoing web hook \cite{SlackOutgoingSlack}. The TextBlob Botlistener listens to port 8080 by default.

\section{Speech Recognizer}
The Speech recognizer, is based on the Google Speech API \cite{GoogleCloudPlatformSpeechPlatform} and is programmed in Csharp. The code is mostly just adapted to post to a web interface, in addition to printing to the console / cli. The Speech Recognizer listens to the detected microphone on the computer by default. It keeps a python dictionary of the users, and their, movie and their most recent sentiment score in memory.

\section{Omdbsearcher}
The Omdbsearcher is based on the Open Movie Data base API.\cite{OMDbDatabase}. When it gets text it will try to search for this text on the OMDB Api and return the JSON formatted result if it finds something, or return a negative result if it does not find anything. The web service listens to port 3000 by default.

\newpage
\section{Architecture}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{system_flow.png}
  \caption{System Flow for Filmchat}
  \label{fig:system_flow}
\end{figure}
The thought of using both of these ways of integrating, was to be able to understand natural language surrounding how people talk and chat about movies, and then later on have the different parts use the same components. This is also why I opted for a modular approach building small components, that I could add to on a later point, and that was easy to test using Postman.

\section{Description of the vision}
The vision of the project, not the actual implementation sadly, aimed to follow the design guidelines give in the chapter about Social Natural User Interfaces \(NUI\) that Wigdor and Wixon talks about in their Brave NUI World book. \cite{WigdorBraveWorld}

Some of the ones I believe fitting to Filmchat are:
\begin{itemize}
\item Create an experience that comes alive with several users, so that the experience
is more fun or efficient when many hands are working simultaneously
\item Enable new users to join, so that the approval of additional users allows them
to easily engage with your application, without disrupting other users already
present.
\item Be able to continue with fewer users, allowing one user to leave without disrupting all others’ experience.
\end{itemize}
I believe them fitting because in the social context of watching a movie the, the experience would come alive, when people are sitting and discussing their likes, and dislikes.
This could, of course be the case in a chat-room setting such as Slack, as well. In Fig\ref{fig:system_flow} the modular approach can be seen, and building more components, or hooking into third party API's could easily be done. 
%----------------------------------------------------------------------------------------
% CHAPTER 3 Interactive System Design
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Interactive System: Design}
\section{System Flow}
The system has three main user inputs \(see Fig\ref{fig:system_flow}\) of which two of them are implemented. They are the Speech recognition, and the Chat. The Chat in the implementation listens to a Slack channel using the outgoing web hook \cite{SlackOutgoingSlack}. It then forwards to the sentiment analysis to extract sentiment and noun phrase \cite{WikipediaNounPhrase} this could have been forwarded to other more advanced third party Natural Language Processors\cite{Wikipedia2017NaturalProcessing} such as Wit.AI\cite{Wit.aiWit.ai} and API.AI\cite{API.AIConversationalAPI.AI}. But I decided to use a light weight Python based for faster development, and easier debugging.

The last input, would have used the same modules, but would have gotten their input from a chat room. The difference would have been the more asynchronous behavior of that kind of interaction, but sentiments and information could have been harvested in much the same way.
\section{Design rationale}

\section{Cost / Benefit of selected interaction techniques}
\subsection{Chat}
Chat is in my view has become a natural way of interacting, several times a day, I interact through chat, be it privately and also work related. A non-exhaustive list of some of the chat interaction I use during a day, or interactions close to chat, are:
\begin{itemize}
\item SMS or Textmessaging
\item Microsoft Teams
\item Slack
\item Lync for Business (formerly Skype)
\item Facebook Chat
\item WhatsApp
\item Snapchat 
\item Tinder
\item Happn
\item Chat Support on company web pages
\end{itemize}

I would also claim that I am in no way unique using many different chat apps throughout the day, and would also therefore claim that chat has become a natural way of interacting with other people.

I would also claim that Chat has a low threshold, lower than using the phone, and it is more instant than using mail. As mail, is in a lot of ways an asynchronous way of communicating.

\subsection{Speech Recognition}
I rarely use speech recognition, in other ways than as a "party trick". To me this is partly because of the social stigma talking to no one out and about, but the other important part is the lacking precision when I have tried it.

In this project I have used the examples from Google, for their Google Speech API
\cite{GoogleCloudPlatformSpeechPlatform}
I have altered the code of the Recognize listener to post the results to a Web API running locally.

But the actual recognized words, has been very limited, typically it has been limited to one, or two words, and not whole sentences.
I have tried both using the laptops integrated microphone, but also a couple of different headsets. None of them have yielded very impressive results.
Caveat: It could be that the actual implemtation example from Google is not the best one there is for my use case, and that improving on the recording part of it, and streaming data would be better. However, spending time working on this has not been a priority, so the actual implementation does nothing more than try to recognize a movie and then post the results about it to a channel on Slack.

%----------------------------------------------------------------------------------------
% CHAPTER 4 Interactive System: Development
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Interactive System: Development}
\section{Overarching Goal}
The main goal for me, was to see what was possible to do, using mostly "off the shelf" components, being libraries, API's or other available premade solutions. This because the project had a limited time, and I wanted to be as effective as possible.
\section{Speech recognition}
I chose to start with an implementation of Speech Recognition, as this was the initial idea of the project.
Some quick googling, and I landed on Google's Speech API \cite{GoogleCloudPlatformSpeechPlatform}. There is an on-line tester, where you can talk, and the Web Page will print what you say.

First I had to enable the Speech API in the Google Cloud console. The API is limited to a certain amount of seconds in the free tier.

First I tried using the Node JS library, but I was not able to get the sound recorder to work properly on my Windows 10 laptop. (It might work better on a linux or Mac, but I am not sure of that)

I then moved on to C\# as I am somewhat familiar with that language. The speech recorder part of that worked much better, and I was able to get the example code to recognize my speech quite quickly.
After I got it working, I modified the code to post the recognized speech to the designated Slack channel.

I selected Slack, as I have developed a couple of things for posting to its interface earlier, and the familiarty meant I was able to test and implement the speech recognition quicker, than had I selected Facebook Messenger API, or Microsofts similar products.
The selection here was then based on familiarity and not based on necessarily "Best of breed" for my specific task.

\section{Omdbsearcher}
Next part of the development was the "omdbsearcher". Here I selected OMDB (Open Movie Database) and had to donate a couple of dollars to get access to their API. There are implementations for IMDB, but they are not, to my knowledge authorized by IMDB, and I was not interested in spending time on a non official api.

IMDB Also has some of its data dumped to a FTP on regular intervals, so creating an own database could have been a future work part for this project. \cite{IMDbAlternativeInterfaces}

The Omdbsearcher I implemented in Node JS, as this is the language I can program in fastest. I tested a couple of request with Postman, which is a tool to create web requests to different endpoints. Postman also has a nice feature where you can get the code for replicating the request in different languages.

I then created a local webserver serving the incoming recognized speech, and then trying to search for them on OMDB. If it got a hit, it posted it to the Slack Channel I had selected.

\section{Textblob Botlistener}
I then checked out a couple of different ways of intepreting language, but have from other projects, already knowledge that Python has such libraries that are good, and with good performance. I therefore winded up with TextBlob, which is a library that can both get sentiment, and noun phrases\cite{WikipediaNounPhrase} from sentences. \cite{TextBlobTutorial:Documentation} 
I therefore just had to create a webservice interface for the TextBlob, and was able to do this quickly.

Description of how it was developed
o If coding project, then include key code snippets and discuss them
o If using a prototype tool, step-by-step description of how the interactive system was created

\section{Spelling corrector}
While the Textblob Botlistener was able to get the noun phrase\cite{WikipediaNounPhrase} of sentences about movie, and that this sometimes is an actual movie, this is not always the case. Therefore I thought about using some autocorrect or spelling corrections, but using the freely available IMDB datasets as valid words \cite{IMDbAlternativeInterfaces}. I found a good example by Peter Norvig \cite{NorvigHowCorrector} on how to do it.
But both the development, and the performance was rather slow, and the dataset would have needed quite a lot of cleaning before I could have used it. Therefore I decided on not to implement this part, and to put it off to a possible future work.

%----------------------------------------------------------------------------------------
% CHAPTER 5 Interactive System: Evaluation
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Interactive System: Evaluation}
• Evaluation plan for the interactive system incorporating at least 5 design principles from the course
Unfortunately, even though the TextBlob (reference) python Library is good, it is still lacking with regards to the goal of my project.
It is able to, sometimes recognize the noun phrase and sentiment, but it is not always the case.
Also keeping context of a ongoing discussion is not something I have looked into at all.
However, I do not think this will not be possible to solve in the future, however they might be "hard" problems.

The speech recognition part of the project, albeit fun, did not fulfill my hopes for what it was able to do. For me to be able to do that, I would probably have to work some time on the actual recording of the audio, and how to optimize this part. Also having the audio recorder being able to switch on and off based on speech patterns.
Also recognizing who actually speaks, is not something I have seen covered in the documentation.
For recognizing users, a chatroom with identified users, is easier.

%----------------------------------------------------------------------------------------
% CHAPTER 6 Ten Meaningful Screens
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Ten Meaningful Screens}
\section{Speech recognition}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{speech_recognition_c-sharp.PNG}
  \caption{Speech recognition using the CSharp SDK for Google Speech}
  \label{fig:c-sharp-speech}
\end{figure}
To get to this screen Fig\ref{fig:c-sharp-speech}, you have to compile the C\# project located under speech-recognizer folder in the GitHub project. \cite{Github2017GitHubProject}
Then you would have to enable the Speech API in the Google Cloud Console, and then finally compile the C\# and run it.
When you are able to execute the program, you type in "Recognize.exe listen". The program will then, using the microphone it detects on your system. Listen for audio and send this to the Speech API for recognition. It will then return them to the cli, and also post them to the Omdbsearcher.

\newpage
\section{Using the Omdbsearcher}
The omdbsearcher Fig\ref{fig:omdbsearcher} part of Filmchat is located under the Omdsearcher folder in the GitHub project \cite{Github2017GitHubProject}. It is based on NodeJS, so first step is to install the neccessary packages by doing: \"npm install\". Then you start it by executing the command: \"node index.js\". This will in turn start a small webserver running on port 3000.
It listens for JSON-formatted data.
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{omdbsearcher.PNG}
  \caption{Omdbsearcher in the cli}
  \label{fig:omdbsearcher}
\end{figure}
For debugging purposes it outputs the search results from the JSON POST it got.

\newpage
\section{Posting to the Omdbsearcher}
For developing purposes, it is possible to post directly to the Omdbsearcher without using the other components. See Fig\ref{fig:postman-film}

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_film_title.PNG}
  \caption{Postman sending film title}
  \label{fig:postman-film}
\end{figure}

\newpage
\section{Result in Slack}
When either posting with Postman, or using the speech recognition component, the results looks like this in the designated Slack channel. Fig\ref{fig:slack-result}

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{slack_result.PNG}
  \caption{Result from post to Omdbsearcher in Slack}
  \label{fig:slack-result}
\end{figure}

\newpage
\section{Sending sentences to TextBlob}
To start the TextBlob Botlistener, go to the textblob folder in the GitHub project \cite{Github2017GitHubProject}. Install the libararies \"pip install libraries\" and run \"python bot-listener.py\" this will start a webservice running on port 8080, that listens for sentences posted to it.
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{textblob_botlistener.PNG}
  \caption{TextBlob Python library webservice}
  \label{fig:slack-result}
\end{figure}

\newpage
\section{Sending sentences to TextBlob part 2}
To send a sentence to the TextBlob Botlistener, use Postman and format the message as seen in Fig\ref{fig:sending-botlistener}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_sentence_01.PNG}
  \caption{Sending a sentence to TextBlob Botlistener}
  \label{fig:sending-botlistener}
\end{figure}

\newpage
\section{Combining the TextBlob Botlistener and the Omdbsearcher}
To send a sentence to the TextBlob Botlistener, use Postman and format the message as seen in Fig\ref{fig:sending-botlistener}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_sentence_01.PNG}
  \caption{Sending a sentence to TextBlob Botlistener}
  \label{fig:sending-botlistener}
\end{figure}

\section{Combining the TextBlob Botlistener and the Omdbsearcher}
To send a sentence to the TextBlob Botlistener, use Postman and format the message as seen in Fig\ref{fig:sending-botlistener}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_sentence_01.PNG}
  \caption{Sending a sentence to TextBlob Botlistener}
  \label{fig:sending-botlistener}
\end{figure}
\newpage
\section{Combining the TextBlob Botlistener and the Omdbsearcher}
To send a sentence to the TextBlob Botlistener, use Postman and format the message as seen in Fig\ref{fig:sending-botlistener}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_sentence_01.PNG}
  \caption{Sending a sentence to TextBlob Botlistener}
  \label{fig:sending-botlistener}
\end{figure}
\newpage
\section{Combining the TextBlob Botlistener and the Omdbsearcher}
To send a sentence to the TextBlob Botlistener, use Postman and format the message as seen in Fig\ref{fig:sending-botlistener}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{postman_sending_sentence_01.PNG}
  \caption{Sending a sentence to TextBlob Botlistener}
  \label{fig:sending-botlistener}
\end{figure}

\newpage
\section{Everything together, almost}
Here \(Fig\ref{fig:everything-together}\), almost all the components are playing together. I write a chat message in the designated Slack channel, while running both the Omdbsearcher, and the TextBlob botlistener, and the result, along with the sentiment gets posted to the channel.
The only part, not here, is the Speech Recognizer, as I was not able to make it recognize more than a word, here and there.

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{everything_together.PNG}
  \caption{Everything together}
  \label{fig:everything-together}
\end{figure}

%----------------------------------------------------------------------------------------
% CHAPTER 7 Five Actionable Events
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Five Actionable Events}
I was not able to find that many actionable events in my Interactive System, but the ones I did find I cover in this section.
\section{}

%----------------------------------------------------------------------------------------
% CHAPTER 8 Two Valuable Outcomes
%----------------------------------------------------------------------------------------

\chapterimage{head3.png}
\chapter{Two Valuable Outcomes}
Generally speaking I would say that, unfortunately my interactive system does not provide that valuable of outcomes.
However here I will try to describe the two, that I thought came the closest.

\section{Speech recognition of movies}
Speech recognition of movies. Being able to search for, and find ratings for movies without having to use the telephone, could be something I might use.
I also thik that this could be a short term valuable outcome of the system. Especially if there was a way of starting to play the movie automatically, once I was happy with the selection. However for a home theater version of this to work, the description and possibly the ability to play a trailer would have to be done on the Television, and maybe not in a Slack channel.

\section{Detect how people talk about movies}
On the longer term I would say that being able to detect how people talk about movies, and then recommending them movies based on what they might also like i think could be very valuable. The difficulty of course would be the level of Natural Language Processing\cite{Wikipedia2017NaturalProcessing} to make this as seamless as I would want it to be. 
My rather simple implementation of using a Python library \cite{TextBlobTutorial:Documentation} to detect noun phrases, and sentiment values, can work on some of the sentences people can say, and write about movies, but it does not keep context. 
\section{Summary of outcomes}
For both of these outcomes I would, however say, that for me it has been valuable to learn, and try to make the mostly "off-the-shelf" technology work together to create these solutions. And with the amount of time available I am rather happy with what I was able to create.

Future work in this might be interesting, and I do believe that better time, and better knowledge of the API's libraries etcetera would mean a better performing Interactive System.

\printbibliography

\vfill
\end{document}